---
title: "Set up iDEP locally"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
IDEP_VERSION <- 94
DATA_VERSION <- "data104"
LOCAL_PORT <- 3838
```

1. (Optional) Download and install Docker on your workstation

https://docs.docker.com/get-docker/

2. Pull the Docker image

- Download the latest version of the image
```bash
docker pull villegar/idep:latest
```

Alternatively,
```bash
docker pull villegar/idep:R-4.0.5
```

3. Set up global variables and create local directories
```bash
IDEP_PATH=/path/to/idep/idep`r IDEP_VERSION`
IDEP_DATA=/path/to/idep/data/
LOCAL_PORT=`r LOCAL_PORT`

mkdir -p $IDEP_DATA/`r DATA_VERSION`
```

<!-- 4. Download the latest version of iDEP from Github -->
<!-- ```bash -->
<!-- cd $IDEP_PATH -->
<!-- git clone https://github.com/iDEP-SDSU/idep -->
<!-- ``` -->

4. Download and extract the database
```bash
wget https://mft.sdstate.edu/public/file/3Y66fppA0Eym0G41taPtRw/data104.tar.gz -O data104.tar.gz
tar xvzf data104.tar.gz --directory $IDEP_DATA/../
rm data104.tar.gz
```

<!-- ```bash -->
<!-- cd $IDEP_DATA/`r DATA_VERSION` -->
<!-- wget https://sdsu.box.com/shared/static/c24f792ojoikpzu0lkpng8uuf9ychwm7.gz -O pathwayDB.tar.gz -->
<!-- tar xvzf pathwayDB.tar.gz -->
<!-- rm pathwayDB.tar.gz -->
<!-- wget https://sdsu.box.com/shared/static/9v1ao6mwhduvrcx793j3answph9gqnkt.gz -O motif.tar.gz -->
<!-- tar xvzf motif.tar.gz -->
<!-- rm motif.tar.gz -->
<!-- wget https://sdsu.box.com/shared/static/mns0k1uvwtfnsohoc89b984ih36nmnz9.gz -O geneInfo.tar.gz -->
<!-- tar xvzf geneInfo.tar.gz -->
<!-- rm geneInfo.tar.gz -->
<!-- wget https://sdsu.box.com/shared/static/qwpdh36vcisgy1hcmadck8i8ezhvr2fh.gz -O data.go.tar.gz -->
<!-- tar xvzf data.go.tar.gz -->
<!-- rm data.go.tar.gz -->
<!-- wget https://sdsu.box.com/shared/static/sorewt7w6iypmhg2k2xhyi8myeit156o.gz -O convertIDs.db.tar.gz -->
<!-- tar xvzf convertIDs.db.tar.gz -->
<!-- rm convertIDs.db.tar.gz -->

<!-- cd $IDEP_DATA/data_go -->
<!-- wget https://raw.githubusercontent.com/villegar/idep/master/data/STRING11_species.csv -->
<!-- ``` -->

5. Download the latest version of iDEP from GitHub:
```bash
git clone https://github.com/iDEP-SDSU/idep $IDEP_PATH
```

6. Deploy a local container
```bash
docker run -dp $LOCAL_PORT:3838 \
  -v $IDEP_DATA:/srv/data \
  -v $IDEP_PATH/shinyapps:/srv/shiny-server \
  villegar/idep:latest
```

Alternatively,
```bash
docker run -dp $LOCAL_PORT:3838 \
  -v $IDEP_DATA:/srv/data \
  -v $IDEP_PATH/shinyapps:/srv/shiny-server \
  villegar/idep:R-4.0.5
```

<!-- ```bash -->
<!-- LOCAL_PORT=`r LOCAL_PORT` -->
<!-- docker run -dp $LOCAL_PORT:3838 \ -->
<!--   -v $IDEP_DATA:/srv/shiny-server/data \ -->
<!--   -v $IDEP_PATH:/srv/shiny-server/ \ -->
<!--   villegar/idep:latest -->
<!-- ``` -->

7. Open a web browser and navigate to the following URL

http://localhost:`r LOCAL_PORT`/idep`r IDEP_VERSION`
