---
title: "Clustering"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clustering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(idepGolem)
devtools::load_all()
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Make data for pre-processing functions
idep_data <- get_idep_data()

YOUR_DATA_PATH <- "E:/idep_9_24/data/data103/data_go/BcellGSE71176_p53.csv"
YOUR_EXPERIMENT_PATH <- "E:/idep_9_24/data/data103/data_go/BcellGSE71176_p53_sampleInfo.csv"

expression_file <- data.frame(
  datapath = YOUR_DATA_PATH
)
experiment_file <- data.frame(
  datapath = YOUR_EXPERIMENT_PATH
)

load_data <- input_data(
  expression_file = expression_file,
  experiment_file = experiment_file,
  go_button = FALSE,
  demo_data_file = idep_data$demo_data_file,
  demo_metadata_file = idep_data$demo_metadata_file
)

converted <- convert_id(
  query = rownames(load_data$data),
  idep_data = idep_data,
  select_org = "BestMatch"
)

all_gene_info <- gene_info(
  converted = converted,
  select_org = "BestMatch",
  idep_data = idep_data
)

converted_data <- convert_data(
  converted = converted,
  no_id_conversion = FALSE,
  data = load_data$data
)

gene_names <- get_all_gene_names(
  mapped_ids = converted_data$mapped_ids,
  all_gene_info = all_gene_info
)

processed_data <- pre_process(
  data = converted_data$data,
  missing_value = "geneMedian",
  data_file_format = 1,
  low_filter_fpkm = NULL,
  n_min_samples_fpkm = NULL,
  log_transform_fpkm = NULL,
  log_start_fpkm = NULL,
  min_counts = .5,
  n_min_samples_count = 1,
  counts_transform = 1,
  counts_log_start = 4,
  no_fdr = NULL
)
```

Clustering is another visualization topic that builds off the first two tabs of
the iDEP application. The package provides methods to cluster and visualize the
processed data. There are other visualization tools in this package that go
along with the clustering that is performed. There is the option to perform
hierarchical or k-means clustering for the main heatmap which will be covered in
the next section. Using the functions below require to first read and use the
operations in the Load_Data and Pre_Process instructions.

###### Determine Cluster Parameters

The sections below help pick parameters for the clustered heatmap.

#### Gene Range

The main heatmap for the clustering has inputs for the number of clusters in the
case of k-means and the number of genes to plot in the heatmap. There are two
plots that can help determine good inputs for these parameters. The first is the
function `sd_density` and shows the standard deviation distribution of the
processed data. There are a max and min genes parameters that will provide lines
to show the standard deviation of the desired range of genes. This will help 
give an idea of the range of genes to plot in the heatmap.

```{r fig.height=5.5, fig.width=7.5}
sd_density(
  data = processed_data$data,
  n_genes_max = 200,
  n_genes_min = 50
)
```

<br>

#### k-Clusters

To select the number of clusters to use in the case fo k-means, we can use the
iDEP function to create an elbow plot. To use this function, we are first going
to have to subset the data with the `process_heatmap_data` function. The input
parameters for this function can be viewed in the help documentation for the
function. We will store the data matrix with the transformations performed to it
in the object `heatmap_data`. This data frame will be used in the function to
create the scree plot. These two functions are demonstrated below.

For information on how to analyze and interpret the elbow plot, please visit
this link https://bl.ocks.org/rpgove/0060ff3b656618e9136b.

```{r fig.height=5.5, fig.width=7.5}
heatmap_data <- process_heatmap_data(
  data = processed_data$data,
  n_genes_max = 200,
  n_genes_min = 50,
  gene_centering = TRUE,
  gene_normalize = TRUE,
  sample_centering = TRUE,
  sample_normalize = TRUE,
  all_gene_names = gene_names,
  select_gene_id = "symbol"
)

k_means_elbow(heatmap_data = heatmap_data)
```

<br>

###### Main Heatmap

Now that we have chosen the number of genes to plot and the heatmap data has
been centered and standardized appropriately. `cluster_meth` provides two
different methods that can be used for the heatmap. The input options can be
seen in the table below.

Input          Description
------------   -----------------
"1"              Hierarchical clustering (Visit https://www.displayr.com/what-is-hierarchical-clustering/)
"2"              k-Means clustering (Visit https://stanford.edu/~cpiech/cs221/handouts/kmeans.html)

The next input `heatmap_cutoff`, will filter the data to ensure that there are
no outliers that would throw off the color scaling of the heatmap. `sample_info`
is the same experiment file that has been used in the previous steps. This will
control the color bar legend on the heatmap and how the samples are grouped. The
options for this input can be found with the call 
`print(c("Sample_Name", colnames(pre_process$sample_info())))`. There are
3 choices for the distance method to be used. The function call `dist_functions`
give the three methods, with a numeric string input to specify which one to use.
Using the distance method, there are 6 possible inputs for the clustering 
method. Details on these methods are in the table below.

Distance Input   Method Descrpition
---------------- ---------------------
"1"              1 - (Pearson's Correlation Coefficient) (Pearson's Correlation Visit https://www.statisticshowto.com/probability-and-statistics/correlation-coefficient-formula/)
"2"              1 - abs(Pearson's Correlation Coefficient)
"3"              Euclidean distance with the `dist` function in R (Visit https://www.cuemath.com/euclidean-distance-formula/)

Clustering Input   Method Description
-----------------  --------------------
 "average"         The distance between two clusters is the average distance between an observation in one cluster and an observation in the other cluster. (Visit for all methods https://support.minitab.com/en-us/minitab/18/help-and-how-to/modeling-statistics/multivariate/how-to/cluster-observations/methods-and-formulas/linkage-methods/#median)
 "complete"        The distance between two clusters is the maximum distance between an observation in one cluster and an observation in the other cluster. 
 "single"          The distance between two clusters is the minimum distance between an observation in one cluster and an observation in the other cluster.  
 "median"          The distance between two clusters is the median distance between an observation in one cluster and an observation in the other cluster.
 "centroid"        The distance between two clusters is the distance between the cluster centroids or means.
 "mcquitty"        The distance is calculated with a distance matrix. See the link above for formula and details.
 
The input`no_sample_clustering` controls whether the columns of the heatmap are
clustered or not (TRUE/FALSE). `heatmap_color_select` sets the color scale that
is used in the heatmap. A vector of three colors should be inputted as
demonstrated below. `row_dend` is TRUE/FALSE for whether to plot the row
dendogram with the heatmap or not. For k-means clusterings, use `k_clusters` to
specify the number of clusters to use. `re-run` is a shiny input and should be
set to FALSE. An example heatmap can be seen below, and is stored in the object
`ht`.

```{r fig.height=5.5, fig.width=7.5}
ht <- heatmap_main(
  data = heatmap_data,
  cluster_meth = "1",
  heatmap_cutoff = 4,
  sample_info = load_data$sample_info,
  select_factors_heatmap = "Sample_Name",
  dist_funs = dist_functions(),
  dist_function = "1",
  hclust_function = "average",
  no_sample_clustering = FALSE,
  heatmap_color_select = c("red", "black", "green"),
  row_dend = TRUE,
  k_clusters = NULL,
  re_run = FALSE
)
```

Due to the interactivity used in the shiny app, the heatmap function shows very
little information. To create an interactive app with the heatmap from above,
use the code below:

```{r eval=FALSE}
InteractiveComplexHeatmap::ht_shiny(ht)
```

<br>

###### Correlation Matrix

Another useful visualization from the iDEP package is a correlation matrix to
show the correaltion between the different samples in the expression data. It 
shows the relationship between the expression values for the different samples.
The values in the matrix are primarily strong positive correlations. This is
because for the vast majority of the genes the expression values are very
similar. The color scale in the correlation plot helps illustrate the slight
differences in correlation and what samples have the strongest relationship. The
function `cor_plot` will create this plot. The `data` input is the entire 
processed datat matrix. `label_pcc` as TRUE will paste the correlation values in
each cell of the correlation heatmap. `heat_cols` is a vector of three colors 
to use for the heatmap color scale. `text_col` controls the color of the 
correlation labels. The code block below demonstrates this function.

```{r fig.height=5.5, fig.width=7.5}
cor_plot(
  data = processed_data$data,
  label_pcc = TRUE,
  heat_cols = c("red", "black", "green"),
  text_col = "white"
)
```

<br>

###### Sample Tree

Another clustering feature is the function to draw a dendogram to show the
clustering of the samples. This will give an idea of how closely the expression
data of each sample is linked to another sample. The clustering is hierarchical
and uses the same functions that were described for the heatmap above. We can
perform centering and normalizing across genes and samples with the 
corresponding input parameters. `hcluster_functions` stores the different 
linkage methods that can be used with the dendogram. The method is specified 
differently than the heatmap above. `hcluster_functions()` is filled into
`hclust_funs`, and the desired method is inputted in `hcluster_function`. The
input for the distance method is the same as the heatmap and is explained above.
The chunk below shows an example call.

```{r fig.height=5.5, fig.width=7.5}
draw_sample_tree(
  tree_data = processed_data$data,
  gene_centering = TRUE,
  gene_normalize = FALSE,
  sample_centering = FALSE,
  sample_normalize = FALSE,
  hclust_funs = hcluster_functions(),
  hclust_function = "average",
  dist_funs = dist_functions(),
  dist_function = "1"
)
```

<br>

###### Running Enrichment

It is possible with iDEP to run enrichment analysis with the genes that are
selected in the `process_heatmap_data` function. This process is explained in
the "Enrichment" instruction. If you wish to perform enrichment with the
`heatmap_data`, save that object in your environment and use the functions
described in that document.

###### Conclusion

This document covered the function used in the Clustering tab of the iDEP
webpage. It demonstrated many visualization techniques, but is in no way
exhaustive of the methods that can be used with the loaded data. For 
troubleshooting, all functions have documentation and the code is available on
Github. 





