---
title: "Load Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Load_Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
#library(idepGolem)
devtools::load_all()
```

This instruction will document how to Load Data and create the data objects that
will be used in the rest of iDEP's functions. This corresponds to the functions
that are utilized in the Load Data tab of the iDEP webpage. The step-by-step
will guide you through loading and running the initial functions to set up the
rest of iDEP analysis.

#### Initial Data Paths

To begin working with iDEP the first step is to identify your working directory
and set it to the path where your data is located. The first two files that will
be worked with is the expression data file and the file containing information
on the experiment design and groupings for the samples. The design for the
expression data file should have the gene IDs as the first column. The ensuing
columns should each correspond to a sample, with the name of the sample in the
first row. The data should then be then correspond (row, column) as the count
for (gene, sample). If an experiment data file is included for the data, the
structure should be similar to the expression data file. The first column should
be the name of an experiment effect. The ensuing column then label each specific
sample. The data will then correspond (row, column) as the group for a
(effect, sample). This instruction will use the built-in demo expression and
experiment files. The first code block will set the data paths to these files
which will be filled into the first functions. We will also run the 
`get_idep_data` function to return a list of the data objects used in the iDEP
data analysis.

```{r Create Data Paths}
idep_data <- get_idep_data()

YOUR_DATA_PATH <- "E:/idep_9_24/data/data103/data_go/BcellGSE71176_p53.csv"
YOUR_EXPERIMENT_PATH <- "E:/idep_9_24/data/data103/data_go/BcellGSE71176_p53_sampleInfo.csv"

expression_file <- data.frame(
  datapath = YOUR_DATA_PATH
)
experiment_file <- data.frame(
  datapath = YOUR_EXPERIMENT_PATH
)
```

#### **`input_data`** Function

This function will read the data from the given data paths and return a list
with the data objects. `data` will be the data matrix of expression data that
was detailed in the paragraph above. `data` will have row names as the gene IDs
and column names matching the names of the samples that each column represents.
`sample_info` will be the experiment design information that was detailed above.
The structure will be the transpose of the original file structure, with the the
rownames now being the column names of the original file. The column names of 
this data matrix will label an effect in the experiment. For example, in the
demo data `sample_info`, the column names are ("p53", "Treatment") for the muted
gene and radiation treatment. If there is no experiment file, the function will
only return the expression data.

For this example, we will store the data in a list called `load_data`.

```{r Load Data}
load_data <- input_data(
  expression_file = expression_file,
  experiment_file = experiment_file,
  go_button = FALSE,
  demo_data_file = idep_data$demo_data_file,
  demo_metadata_file = idep_data$demo_metadata_file
)
```

#### iDEP Database Functions

The next step of process is to use the gene IDs from the expression data to get
the gene information and determine the species that the expression samples are
from using the iDEP database. The following functions will also convert the IDs,
if they are recognized, to ensembl and symbol name. These functions have an
input parameter to specify the species that the data is for. The line below will
shows how to search iDEP to find the numeric code for the species you wish to
use. To allow iDEP to find the species, use `"BestMatch"`. To use a numeric code
for a specfic species, make sure to use string quotes, for example `"107"`.

```{r}
# Species search for input data
idep_data$species_choice[grep("Mouse", names(idep_data$species_choice))]
```

Each iDEP database function is outlined below.

1. **`convert_id`**

    This function will take in a 'query' and use the 'idep_data' to match the 
    provided gene IDs with the corresponding ensembl ID. The IDs from the
    inputted data are currently the rownames of the data frame `data` stored in
    the list `load_data`. `select_org` is a parameter to provide the species
    that the expression data is from. `convert_id` returns a list containing
    information about the query results. The converted IDs are are accessed by
    `converted$ids`. `print(converted$species_matched[1,])` shows the species
    that was recognized by the provided gene IDs.
    
```{r}
converted <- convert_id(
  query = rownames(load_data$data),
  idep_data = idep_data,
  select_org = "BestMatch"
)

print(converted$species_matched[1,])
```
    

2. **`gene_info`**

    Gathering the information on the converted ensembl IDs is the next step. 
    The `converted` results are passed into the `gene_info` function to retrieve
    gene name information that is stored in `idep_data` as well. If `select_org`
    is `"BestMatch"`, the function will use the matched species from converted
    to retrieve the gene data. If a species code is provided, it will use that
    species directly. `gene_info` will then read a gene data file from the
    `idep_data` and return the resulting data frame. This data will be used to
    map to more informative gene names in the coming functions.
    
```{r}
all_gene_info <- gene_info(
  converted = converted,
  select_org = "BestMatch",
  idep_data = idep_data
)
```
    
    
3. **`convert_data`**

    This step will return the original data from `input_data` with the converted
    ensembl IDs as rownames, and a data frame of the mapped IDs. It uses the IDs
    that were matched with `convert_id` to swap the current rownames with the 
    corresponding ensembl ID. `no_id_conversion` can be set to `TRUE` if the 
    desired IDs are the original ones, but the simpler solution is to simply not
    run the function. The returned list contains the data with converted IDs in 
    `data` and IDs in `mapped_ids`.
    
```{r}
converted_data <- convert_data(
  converted = converted,
  no_id_conversion = FALSE,
  data = load_data$data
)
```


4. **`get_all_gene_names`**

    This function combines all the gene name information that has been gathered
    in the preceding functions. `mapped_ids` are the matched IDs from 
    `converted`, and `all_gene_info` is the data read from idep based off the
    identified species. This function returns a data frame with `User_ID` as the
    original IDs, `ensembl_ID` containing the converted IDs, and `symbol`
    containing the informative gene names. This data frame can be used
    throughout the package to matach gene IDs.
    
```{r}
gene_names <- get_all_gene_names(
  mapped_ids = converted_data$mapped_ids,
  all_gene_info = all_gene_info
)
```

#### Conclusion

This concludes the function calls and processes from the first tab of the iDEP
program. There is no real data manipulation that happens, but the objects that
are created are essential for the processes that will happen in the following
instructions. For troubleshooting, all functions have documentation and the code
is available on Github. (https://github.com/gexijin/idepGolem)