---
title: "Pre-Process"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pre_Process}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
#library(idepGolem)
devtools::load_all()
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Make data for pre-processing functions
idep_data <- get_idep_data()

YOUR_DATA_PATH <- "E:/idep_9_24/data/data103/data_go/BcellGSE71176_p53.csv"
YOUR_EXPERIMENT_PATH <- "E:/idep_9_24/data/data103/data_go/BcellGSE71176_p53_sampleInfo.csv"

expression_file <- data.frame(
  datapath = YOUR_DATA_PATH
)
experiment_file <- data.frame(
  datapath = YOUR_EXPERIMENT_PATH
)

load_data <- input_data(
  expression_file = expression_file,
  experiment_file = experiment_file,
  go_button = FALSE,
  demo_data_file = idep_data$demo_data_file,
  demo_metadata_file = idep_data$demo_metadata_file
)

converted <- convert_id(
  query = rownames(load_data$data),
  idep_data = idep_data,
  select_org = "BestMatch"
)

all_gene_info <- gene_info(
  converted = converted,
  select_org = "BestMatch",
  idep_data = idep_data
)

converted_data <- convert_data(
  converted = converted,
  no_id_conversion = FALSE,
  data = load_data$data
)

gene_names <- get_all_gene_names(
  mapped_ids = converted_data$mapped_ids,
  all_gene_info = all_gene_info
)
```

This instruction builds off the objects and functions from the "Load_Data"
vignette. Please see that document before using the functions described below.
We will now transform and process the data based off the type that is being
studied. The package also provides visualization functions that will provide a
foundation for exploratory data analysis and good overview of the data.

#### **`pre_process`** Function

Processing the data requires many options to ensure the data gets to the desired
form. This is the data that will be used in the exploratory analysis to come.
The data to be processed is the `converted_data` from the "Load_Data" vignette.
The parameters `missing_value` and `data_file_format` have a list of input
values to select from for processing. These values and descriptions can be found
in the tables below. The `missing_value` parameter determines how missing data
entries will be filled in, and the `data_file_format` parameter tells the
function what type of data is being processed. Depending on the
`data_file_format` chosen, additional parameters will need to be provided. These
are explained below.

Input (`missing_value`)    Description
-------------------------  -----------------------------------------------------------
`geneMedian`               Substitute median expression value for gene across all samples
`treatAsZero`              Substitute "0" for all missing values
`geneMedianInGroup`        Substitute median expression value for gene within sample group

Input (`data_file_format`)   Description
---------------------------  -----------------------------------------------------------------------
`1`                          Read counts data (recommended)
`2`                          Normalized expression values (RNA-seq FPKM, micro-array, etc.)
`3`                          Fold-changes and corrected P values from CuffDiff or any other program

<br>

##### **Read counts data**

For read counts data, the parameters `min_counts`, `n_min_samples_counts`,
`counts_transform` and `counts_log_start` need input values. The parameters 
`min_counts` and `n_min_samples` work to filter the genes out of the data that
do not have sufficient counts. The function counts the columns that have more
counts than `min_counts` for each gene and filters out the genes with less
columns than `n_min_samples`. For example, if you only wanted genes with more
than 0 counts for each sample, you would set `min_counts = .5` and 
`n_min_samples = ncol(converted_data)`. The parameter `counts_transform`
specifies the transformation to perform on the counts data. This parameter has 3
input options which are outlined below. If `counts_transform = 1`,
`counts_log_start` determines the constant to add to CPM to ensure that there
are no `log(0)` errors. All other parameters can be set to `NULL`

Input (`counts_transform`)   Description
---------------------------  --------------------------------------------------
`1`                          EdgeR: log2(CPM+c) (Counts per Million + constant) (Visit https://www.rdocumentation.org/packages/SparkR/versions/2.1.2/topics/log2)
`2`                          VST: variance stabilizing transform (Visit https://www.rdocumentation.org/packages/DESeq2/versions/1.12.3/topics/vst)
`3`                          rlog: regularized log (slow) (Visit https://www.rdocumentation.org/packages/DESeq2/versions/1.12.3/topics/rlog)

<br>

##### **Normalized expression values Parameters**

If the data being processed is for normalized expression values, the parameters
`low_filter_fpkm`, `n_min_samples_fpkm`, `log_transform_fpkm` and 
`log_start_fpkm` need input values. The parameters `low_filter_fpkm` and
`n_min_samples_fpkm` work together to filter out the genes that do not have
enough counts. The function will count the columns that are greater than
`low_filter_fpkm` and discard the gene if it is not more than
`n_min_samples_fpkm` columns. The parameter `log_transform_fpkm` is `TRUE/FALSE`
for whether to perform a log transform on the data or not. If
`log_transform = TRUE`, then `log_start_fpkm` is the value that will be added to
ensure there are no `log(0)` errors. All other values can be set to `NULL`.

<br>

##### **Fold-changes and corrected P values**

For fold-changes and corrected p-values, there is only one additional parameter
to provide an input for, `no_fdr`. This is a logical `TRUE/FALSE` to specify
whether the data contains corrected p-values or not. `TRUE` denotes fold-change
only data. All other parameters can be set to `NULL`.

<br>
<br>

The demonstration loaded in the "Load_Data" instruction is read counts data. The
code below shows an example use of the `pre_process` function using the example
data. This data has the converted ensembl IDs. We will return a large list of
the results of the processing from this function. The objects in this list are
defined in the table below.

Output               Description
-------------------  -------------
data                 Data matrix that has the pre-process operations performed to it.
mean_kurtosis        The calculated kurtosis of the mean counts for a gene. (Visit https://www.rdocumentation.org/packages/e1071/versions/1.7-2/topics/kurtosis)
raw_counts           The raw counts before pre-processing.
data_type_warning    Set to "1" if iDEP recognizes the data as different than the inputted `data_file_format`.
data_size            Length 4 vector giving the dimensions of the data before and after processing.
p_vals               For fold-change data has p-values, they are returned here. 

```{r}
processed_data <- pre_process(
  data = converted_data$data,
  missing_value = "geneMedian",
  data_file_format = 1,
  low_filter_fpkm = NULL,
  n_min_samples_fpkm = NULL,
  log_transform_fpkm = NULL,
  log_start_fpkm = NULL,
  min_counts = .5,
  n_min_samples_count = 1,
  counts_transform = 1,
  counts_log_start = 4,
  no_fdr = NULL
)
```

#### Exploratory Data Analysis

Now that the data has been processed, we can begin data analysis. It is possible
to create unique and informative plots with the processed data, and iDEP has
built in functions to assist with that. These plots can then also be saved in
an object and have additional ggplot features added to them.

<br>

#### **`total_counts_ggplot`** Function

This function will only work for read counts data, so if that is not the format
of the processed data, please skip this function. For read counts data, this
function creates a barplot to show the counts in millions for each sample in the
data. `counts_data` is the raw counts that is returned from the `pre_process`
function. `sample_info` is the experiment design data from the `input_data`
function. 

```{r fig.height=5.5, fig.width=7.5}
total_counts_ggplot(
  counts_data = processed_data$raw_counts,
  sample_info = load_data$sample_info
)
```

<br>

#### **`eda_scatter`** Function

Another useful plot is a scatter plot of the transformed expression for two
samples in the data. `eda_scatter` provides a scatter plot of two specified
sample (columns) of the processed data. `plot_xaxis` and `plot_yaxis` are
string names of the columns to be plotted. The options for these inputs can be
seen using `colnames(processed_data$data)`. `processed_data` is the returned
data from the `pre_process` function. 

```{r fig.height=5.5, fig.width=7.5}
colnames(processed_data$data)

eda_scatter(
  processed_data = processed_data$data,
  plot_xaxis = "p53_IR_1",
  plot_yaxis = "p53_mock_1"
)
```

<br>

#### **`eda_density`** Function

This plotting function provides a plot of the distribution of counts values for
each sample in the data. The information is displayed in a density ridge plot in
the ggplot format. Again, the data provided is returned from the `pre_process`
function. The `sample_info` provides information on each sample in the dataset.

```{r fig.height=5.5, fig.width=7.5}
eda_density(
  processed_data = processed_data$data,
  sample_info = load_data$sample_info
)
```

<br>

#### `merge_data` Function

The current data is using the converted ensembl IDs as rownames in a large data
matrix, but the following function will create a data frame containing the gene
names from the `get_all_gene_names` function. `merge_data` takes the
`gene_names` data frame and a data matrix with ID rownames and creates a data
frame with columns for each ID that idep has matched from the original IDs.
This is useful to view different IDs for the same gene. It also enables writing
CSV files that can be saved and be used to reproduce results from iDEP. The
following code will merge the gene names with the raw counts data and the
transformed data from `pre_process`.

```{r}
merged_raw_counts <- merge_data(
  all_gene_names = gene_names,
  data = processed_data$raw_counts,
  merge_ID = "ensembl_ID"
)

merged_data <- merge_data(
  all_gene_names = gene_names,
  data = processed_data$data,
  merge_ID = "ensembl_ID"
)
```

<br>

#### Plotting Individual Genes

iDEP also has a function that visualizes individual genes and their expression
behavior across the samples. It is possible to determine the genes to plot by
looking at the merged or processed data. It is also possible to use the 
`gene_names` data frame to search for a gene ID. The code below demonstrates how
to search for a gene ID.

```{r}
gene_names[grep("H4c1", gene_names$symbol), ]
```

Depending on the type of gene ID that is desired for the plot, the function
below will swap the gene IDs with the processed data matrix. This new data
matrix will be used in the individual plot function using the selected ID type
and the desired gene. The code below shows how to swap the gene IDs to the
symbol name.

```{r}
ind_plot_data <- rowname_id_swap(
  data_matrix = processed_data$data,
  all_gene_names = gene_names,
  select_gene_id = "symbol"
)
```

The function `individual_plots` will create the plot for the desired gene. We
will use the data that we created in the code above for `individual_data`. If
the data comes with an experiment file, the sample information that was loaded
in the first instruction will be used in `sample_info`. The gene to be plotted
can be specified in the `selected_gene` parameter. Multiple genes can be
inputted as a vector to plot them at the same time and compare. There are two
different types of plots that can be outputted with this function. The parameter
`gene_plot_box` controls which plot will be outputted. Setting this input to
TRUE will output a lineplot that gives the expression across all samples. FALSE
will output a bar plot that gives the mean expression across the groups detected
in the sample names. For the barplot, `use_sd` as TRUE will give a standard
deviation bar instead of a standard error bar. `lab_rotate` controls the
rotation of the x-axis labels. The outputted object is a formatted ggplot.

```{r fig.height=5.5, fig.width=7.5}
individual_plots(
  individual_data = ind_plot_data,
  sample_info = load_data$sample_info,
  selected_gene = "H4c1",
  gene_plot_box = FALSE,
  use_sd = FALSE,
  lab_rotate = 90
)
```

<br>

#### Conclusion

This concludes all the functions utilized in the Pre-Process portion of the iDEP
package. The data can be plotted in many different ways, but hopefully the
functions detailed above provide a good starting point. For troubleshooting, all
functions have documentation and the code is available on Github.
(https://github.com/gexijin/idepGolem)