## 2 PreProcessing Result

Before analysis the data, we need some preprocess procedure to make the data tidy and clean.

TODO: more detail about preprocess


### 2.1 Preprocess parameters

All preprocess setup are listed here.

** Common Preprocess Setting **

[Missing values imputation](link to Missing values imputation): `r input$selectMissingValueImputationMethod`

[Do not convert gene IDs to Ensembl](link to isNoIDConversion): `r input$isNoIDConversion`


```{r, child = 'PreProcess/ReadCountSettings.Rmd', eval=(input$dataFileFormat == 1)}
```

```{r, child = 'PreProcess/ReadCountSettings_PseudoC.Rmd', eval=(input$dataFileFormat == 1 && input$selectCountsTransform == 1)}
```

```{r, child = 'PreProcess/FPKMSettings.Rmd', eval=(input$dataFileFormat == 2)}
```

### 2.2 Guess Species

TODO: Some of description about guess species

The species is `r PreProcessCtrl$GetGuessSpeciesResult(ConvertedIDResult)`

### 2.3 Total Read Count (millions)

```{r}
PreProcessCtrl$GetTotalReadCountsPlot(PreProcessResult$rawCount)
```

### 2.4 Distribution of transformed data

```{r}
PreProcessCtrl$GetTransformedDataBoxPlot(PreProcessResult$dat)
```

### 2.5 Density plot of transformed data

```{r}
PreProcessCtrl$GetTransformedDataDensityPlot(PreProcessResult$dat)
```

### 2.6 Scatter plot of first two samples

```{r}
PreProcessCtrl$GetTransformedDataScatterPlot(PreProcessResult$dat)
```



