## PreProcessing Result

Before analysis the data, we need some preprocess procedure to make the data tidy and clean.

TODO: more detail about preprocess


### Preprocess parameters

All preprocess setup are listed here.
*** Common Preprocess Setting ***
[link to Missing values imputation](Missing values imputation): `r input$selectMissingValueImputationMethod`
[link to isNoIDConversion](Do not convert gene IDs to Ensembl): `r input$isNoIDConversion`


```{asis, echo=(input$dataFileFormat == 1)}
*** Readcount only settings ***
[link to numMinCounts](Min. CPM): `r input$numMinCounts`
[link to numNMinSamplesInCountCase](n libraries): `r input$numNMinSamplesInCountCase`
[link to selectCountsTransform](Transform algorithm): `r switch(input$selectCountsTransform, 'EdgeR', 'VST', 'rlog')`
```
```{asis, echo=(input$dataFileFormat == 1 && input$selectCountsTransform == 1)}
[link to Pseudo count c](Pseudo count c): `r input$numCountsLogStart`
```

```{asis, echo=(input$dataFileFormat == 2)}
*** FPKM Data Related Settings ***
[link to numMinFPKM](Min. level): `r input$numMinFPKM`
[link to numNMinSampleInFPKMCase](n samples): `r input$numNMinSampleInFPKMCase`
[link to isApplyLogTransFPKM](Log Transformation): `r input$isApplyLogTransFPKM`
[link to numFPKMLogStart](Constant c for started log): `r input$numFPKMLogStart`
```

### Guess Species

TODO: Some of description about guess species

The species is `r print(PreProcessCtrl$GetGuessSpeciesResult(ConvertedIDResult))`

### Total Read Count (millions)

```{r}
PreProcessCtrl$GetTotalReadCountsPlot(PreProcessResult$rawCount)
```

### Distribution of transformed data

```{r}
PreProcessCtrl$GetTransformedDataBoxPlot(PreProcessResult$dat)
```

### Density plot of transformed data

```{r}
PreProcessCtrl$GetTransformedDataDensityPlot(PreProcessResult$dat)
```

### Scatter plot of first two samples

```{r}
PreProcessCtrl$GetTransformedDataScatterPlot(PreProcessResult$dat)
```



